---
title: "anomhot_2018_2019_ca"
author: "Brittany Shea"
date: "`r Sys.Date()`"
output: html_document
---

# Download temperature data: https://github.com/rmp15/PRISM-grids-into-FIPS-ZIP-censustract-USA/tree/main/output/fips/tmean

# Load libraries
```{r}
library(readr)
library(viridis)
library(stringr)
library(tidyverse)
library(lubridate)
library(rgdal)
library(ggplot2)
library(raster)
library(sp)
library(plyr)
library(graticule)
library(zoo)
library(dplyr)
library(conflicted)
```

# Remove conflicts when working with dplyr & spatial data
```{r}
conflict_prefer("filter", "dplyr")
```

# Load temperature data 1981-2020
```{r}
dat1 = readRDS("data/weighted_area_raster_fips_tmean_daily_1981.rds")
dat2 = readRDS("data/weighted_area_raster_fips_tmean_daily_1982.rds")
dat3 = readRDS("data/weighted_area_raster_fips_tmean_daily_1983.rds")
dat4 = readRDS("data/weighted_area_raster_fips_tmean_daily_1984.rds")
dat5 = readRDS("data/weighted_area_raster_fips_tmean_daily_1985.rds")
dat6 = readRDS("data/weighted_area_raster_fips_tmean_daily_1986.rds")
dat7 = readRDS("data/weighted_area_raster_fips_tmean_daily_1987.rds")
dat8 = readRDS("data/weighted_area_raster_fips_tmean_daily_1988.rds")
dat9 = readRDS("data/weighted_area_raster_fips_tmean_daily_1989.rds")
dat10 = readRDS("data/weighted_area_raster_fips_tmean_daily_1990.rds")
dat11 = readRDS("data/weighted_area_raster_fips_tmean_daily_1991.rds")
dat12 = readRDS("data/weighted_area_raster_fips_tmean_daily_1992.rds")
dat13 = readRDS("data/weighted_area_raster_fips_tmean_daily_1993.rds")
dat14 = readRDS("data/weighted_area_raster_fips_tmean_daily_1994.rds")
dat15 = readRDS("data/weighted_area_raster_fips_tmean_daily_1995.rds")
dat16 = readRDS("data/weighted_area_raster_fips_tmean_daily_1996.rds")
dat17 = readRDS("data/weighted_area_raster_fips_tmean_daily_1997.rds")
dat18 = readRDS("data/weighted_area_raster_fips_tmean_daily_1998.rds")
dat19 = readRDS("data/weighted_area_raster_fips_tmean_daily_1999.rds")
dat20 = readRDS("data/weighted_area_raster_fips_tmean_daily_2000.rds")
dat21 = readRDS("data/weighted_area_raster_fips_tmean_daily_2001.rds")
dat22 = readRDS("data/weighted_area_raster_fips_tmean_daily_2002.rds")
dat23 = readRDS("data/weighted_area_raster_fips_tmean_daily_2003.rds")
dat24 = readRDS("data/weighted_area_raster_fips_tmean_daily_2004.rds")
dat25 = readRDS("data/weighted_area_raster_fips_tmean_daily_2005.rds")
dat26 = readRDS("data/weighted_area_raster_fips_tmean_daily_2006.rds")
dat27 = readRDS("data/weighted_area_raster_fips_tmean_daily_2007.rds")
dat28 = readRDS("data/weighted_area_raster_fips_tmean_daily_2008.rds")
dat29 = readRDS("data/weighted_area_raster_fips_tmean_daily_2009.rds")
dat30 = readRDS("data/weighted_area_raster_fips_tmean_daily_2010.rds")
dat31 = readRDS("data/weighted_area_raster_fips_tmean_daily_2018.rds")
dat32 = readRDS("data/weighted_area_raster_fips_tmean_daily_2019.rds")
```

# Combine 1981-2010 data frames into one data frame
```{r}
bind_weighted_area_raster_fips_tmean_daily_1981_2010 <- rbind(dat1, dat2, dat3, dat4, dat5, dat6, dat7,
                                                              dat8, dat9, dat10, dat11, dat12, dat13, dat14,
                                                              dat15, dat16, dat17, dat18, dat19, dat20, dat21,
                                                              dat22, dat23, dat24, dat25, dat26, dat27, dat28,
                                                              dat29, dat30)
```

# Combine 2018-2019 data frames into one data frame
```{r}
bind_weighted_area_raster_fips_tmean_daily_2018_2019 <- rbind(dat31, dat32)
```

# Change format of date columns
```{r}
bind_weighted_area_raster_fips_tmean_daily_1981_2010 <- bind_weighted_area_raster_fips_tmean_daily_1981_2010 %>%
  dplyr::mutate(date = lubridate::dmy(date))

bind_weighted_area_raster_fips_tmean_daily_2018_2019 <- bind_weighted_area_raster_fips_tmean_daily_2018_2019 %>%
  dplyr::mutate(date = lubridate::dmy(date))
```

# Group data frames by week in year
```{r}
bind_weighted_area_raster_fips_tmean_daily_1981_2010$week_num <- strftime(bind_weighted_area_raster_fips_tmean_daily_1981_2010$date, format = "%V")

bind_weighted_area_raster_fips_tmean_daily_2018_2019$week_num <- strftime(bind_weighted_area_raster_fips_tmean_daily_2018_2019$date, format = "%V")
```

# For months with 53 values, average it with 52
```{r}
bind_weighted_area_raster_fips_tmean_daily_1981_2010$week_num[which(bind_weighted_area_raster_fips_tmean_daily_1981_2010$week_num == "53")] = "52"

bind_weighted_area_raster_fips_tmean_daily_2018_2019$week_num[which(bind_weighted_area_raster_fips_tmean_daily_2018_2019$week_num == "53")] = "52"
```

# Group "bind_weighted_area_raster_fips_tmean_daily_1981_2010" by week and county, then calculate 85th percentile using weekly long-term average
```{r}
dname = 'hot_85th'

fips85thpercentile = bind_weighted_area_raster_fips_tmean_daily_1981_2010 %>% 
  group_by(week_num,fips) %>%
  dplyr::summarise(hot_85th = quantile(tmean, probs = 0.85)) %>%
  dplyr::mutate(fips = as.numeric(as.character(fips)))
```

# Group "bind_weighted_area_raster_fips_tmean_daily_2018_2019" by week and county, then find count of number of days over a threshold (>24°C)
```{r}
dname = 'tmean'

threshold = 24

over24_tmean_daily_2018_2019 <- bind_weighted_area_raster_fips_tmean_daily_2018_2019 %>%
  dplyr::mutate(over_24 = ifelse(get(dname) > threshold,1,0)) %>%
  dplyr::group_by(fips) %>%
  dplyr::mutate(fips = as.numeric(as.character(fips)))
```

# Perform left join
```{r}
join_tmean_daily_over24_85th_2018_2019 <- over24_tmean_daily_2018_2019 %>%
  left_join(fips85thpercentile, by = c("week_num" = "week_num", "fips" = "fips"))
```

# Create new column to determine whether tmean_daily meets thresholds
```{r}
#first find days that meet 85th percentile threshold and the median for the tmean column
join_tmean_daily_85th_2018_2019 <- join_tmean_daily_over24_85th_2018_2019 %>%
  dplyr::mutate(anomwarm = ifelse(tmean >= hot_85th, 1,0)) %>% 
    filter(anomwarm == 1) %>% 
    subset(fips %in% 6001:6115)

#find the median for the tmean column
median(join_tmean_daily_85th_2018_2019$tmean)

#find the 25th percentile for the tmean column
quantile(join_tmean_daily_85th_2018_2019$tmean, probs = 0.25)

#find the 75th percentile for the tmean column
quantile(join_tmean_daily_85th_2018_2019$tmean, probs = 0.75)


#Create new column to determine whether tmean_daily meets 2 thresholds
join_tmean_daily_over24_85th_2018_2019 <- join_tmean_daily_over24_85th_2018_2019 %>%
  dplyr::mutate(anomwarm = ifelse(over_24 == 1 & tmean >= hot_85th, 1,0))
```

# Find number of hot days by county in CA only and meets threshold
```{r}
join_tmean_daily_over24_85th_2018_2019 <- join_tmean_daily_over24_85th_2018_2019 %>%
  filter(anomwarm == 1) %>% 
  subset(fips %in% 6001:6115)

aw2join = join_tmean_daily_over24_85th_2018_2019

ncol = paste(aw2join$fips,aw2join$date,sep = "")

aw2join$fipdate = ncol

aw2join2 = aw2join[ ,c("fipdate","date","fips","anomwarm")]

colnames(aw2join2) = c("fipdate", "date_aw",  "fips_aw", "exposed_aw")

write_csv(aw2join2, "output/aw2join.csv")


```

# Check data with a table
```{r}
table(join_tmean_daily_over24_85th_2018_2019$anomwarm,join_tmean_daily_over24_85th_2018_2019$month)
```

# Find number of warm days by county over time
```{r}
county_count_warm <- join_tmean_daily_over24_85th_2018_2019 %>% 
  group_by(fips) %>%
  dplyr::summarize(total_warm = sum(anomwarm))
```

# Save file
```{r}
write_csv(county_count_warm, "output/aw_data.csv")

```